The development of children's ability to track and predict turn structure in conversation

### 2015-11-28 paper/CasillasFrank-Anticip.tex
Small changes to randrun analysis/plots; paper updates in progress

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See an example run-though of the videos here: www.youtube.com/channel/UCGEZQcM9t8Zfjqqi\_B1Q5Sw.}

@@

\subsection{Results and discussion}
\label{sec:results1}

@@

\begin{table}[t]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\ 
    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\ 
    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\ 
    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\ 
    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\ 
    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\ 
    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\ 
    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\ 
%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
\label{tab:e1_look}
\end{table}
%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
%\end{center}
%\caption{Proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
%\label{fig:randvsrealFL}
%\end{figure}
%
%To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition. Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., non-native prosodic cues and non-verbal signals). For all further analyses we then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
%
%Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09). But the tendency to make more anticipatory switches for English than non-English videos disappeared with age (Figure \ref{fig:conditionsFL}). While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults. Otherwise, age-based differences in predictive gaze shifts were minor.
%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
%\end{center}
%\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
%\label{fig:conditionsFL}
%\end{figure}
%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
%\end{center}
%\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
%\label{fig:questionsFL}
%\end{figure}
%
%Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}). Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos.
%

@@

\begin{table}[t]
\begin{center}
  \begin{tabular}{lcccc}
  \textbf{\textit{Children}} &&&& \\
    \hline
                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
    \hline
(Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
LgGroupNE &          -3.29347 &   0.96045 & -3.429 & 0.000606 \\
TypeS          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  \\
Age:LgGroupNE &       0.52065 &   0.21190 &  2.457 & 0.014008 \\  
Age:TypeS          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
LgGroupNE:TypeS &     2.68166 &   1.35016 &  1.986 & 0.047013 \\  
Age:LgGroupNE:TypeS & -0.45632 &   0.30163 & -1.513 & 0.130315  \\   
    \hline
  &&&& \\
  \textbf{\textit{Adults}} &&&& \\
    \hline
                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
    \hline
(Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
LgGroupNE    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
TypeS             &        -4.4953 &    1.3139 & -3.421 & 0.000623  \\
Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
LgGroupNE:TypeS &           3.2972 &    1.6101 &  2.048 & 0.040581  \\  
LgGroupNE:Duration &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
TypeS:Duration          &  10.5107 &    3.3459 &  3.141 & 0.001682  \\
LgGroupNE:TypeS:Duration & -6.3156 &    4.4926 & -1.406 0.159790 \\    
    \hline
  \end{tabular}
\end{center}
  \caption{Model output for children and adults' anticipatory gaze switches.}
\label{tab:E1-models}
\end{table}

Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of a language condition-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05) and an age-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05). There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{t}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{t}=-1.27, \textit{p}$=$.2, respectively).

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$300 msec around the inter-turn gap.} 
\label{fig:shuffling}
\end{figure}

\subsubsection{Random baseline comparison}

We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transitions (Figure \ref{fig:shuffling}). This involved: (1) randomizing the order and temporal placement of the analysis windows (Figure \ref{fig:criterion}) for each stimulus, thereby randomly redistributing the windows of analysis across the eye-tracking signal for each stimulus, then (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}) with the randomly permuted analysis windows, and finally (3) entering the resulting anticipatory gazes into the same statistical model we used with the original data (above). Properties of each turn transition (e.g., prior speaker identity, transition type, duration, etc.) stayed constant, even though its onset and offset time within the eye-tracking signal was shuffled.

In effect, this procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior from the same data. The random permutations represent average anticipatory gaze rate over all possible starting points; a random baseline. If we anticipatory gaze rates turned out to the same for the original and randomly permuted versions of the data, we would have no evidence for turn-relevant gaze switching. Further, by running both data sets through the identical statistical models, we can estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition) arose from random looking.

We completed this random baseline procedure on 5,000 permutations of the original turn transition analysis windows and then compared the beta estimates from the original data models (above) to the distribution of beta estimates for the models of the 5,000 randomly permuted datasets. We estimated whether each effect in our original statistical models differed from chance by calculating how many random beta estimates were smaller than the original beta estimate for each predictor, using the absolute value of all beta estimates for a two-tailed test. For example, children's original ``language condition'' beta estimate was $|$-3.29$|$, which is greater than 96.9\% of $|$beta estimates$|$ from the models of randomly-permuted data. This leads us to conclude that that the effect of language condition in the original model is unlikely to be the result of random gaze shifting. We excluded the output of random-permutation models that did not converge to remove unreliable \textit{$\beta$} estimates from our percentile calculations below (22.5\% and 24.4\% of models for children and adults, respectively).

\begin{figure}[h]
  \centering
  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-dist.png}\label{fig:E1-ChiBetas}}
  \hfill
  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-dist.png}\label{fig:E1-AduBetas}}
  \caption{Random-permutation and original $|$beta estimates$|$ for predictors of children and adults' anticipatory gaze rates. Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each beta estimate distribution.}
\end{figure}

Most of the significant predictors from models of the original, turn-related data can not be explained by random looking. The children's data showed strong evidence of differentiation between the random and original beta estimates for the effect of language condition ($>$96.9\% of random beta estimates) and the age-language condition interaction ($>$93.9\%), with somewhat weaker support for the language condition-transition type interaction ($>$85.1\%), and no support for the effect of gap duration ($>$73\%; Figure \ref{fig:E1-ChiBetas}). The adult data showed moderate evidence for the effect of transition type ($>$88.6\%), weak evidence for the interaction of transition type and duration ($>$80.6\%), and no evidence for the interaction of language condition and transition type ($>$59.8\%; Figure \ref{fig:E1-AduBetas}).

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
\end{center}
\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data.} 
\label{fig:E1-randvsreal}
\end{figure}

% Statistical model comparing random to real data at each age.

\subsubsection{Summary}

%Several results from Experiment 1 were unexpected. First, we found no significant effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions). It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were. To find out when children first begin making spontaneous anticipatory looks to responders, it is necessary to recruit younger participants. Experiment 2 thus includes a sample of one- and two-year-olds.

%Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches. An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand. As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn taking closely. To identify which linguistic cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.

%Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax. We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information. Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case. So, although using non-English videos was a natural way to control for lexical information, it did not allow us to assess the role of prosody. Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation. Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language. In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.

%% E1 DISCUSSION LEFTOVER TEXT

%Language did not affect adults' anticipatory looking behavior, possibly because the animated, child-directed prosody and non-verbal cues made it possible for them to follow the interaction and track the current speaker easily.
% As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.
%Contrary to our expectation, there was no effect of age. Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition. We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data. One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
%It is also possible that the timing of anticipatory gaze switches changes as children get older. That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age. For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older. If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition. To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogels2015}. Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn). But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}). The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.
%Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access. There was no effect of transition type alone. Rather, the model confirmed a two-way interaction of linguistic condition and transition type. Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic). Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}). The children's model confirmed a three-way interaction of age, linguistic condition, and transition type. Lexicosyntactic information might then be critical for young children to identify questions in conversation. Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next. Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers. Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.
%The increased switching after questions indicates that, so long as they have native lexical and prosodic cues, children as young as three spontaneously distinguish questions from other types of speech acts in third-party conversation. Children may treat questions differently from non-questions in conversation because they are predictably followed by an immediate response (i.e., the answer). Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), children could even be predicting what type of response they will hear, and look at the responder for confirmation. In contrast, non-questions do not usually require an immediate response, so it is not as easy to predict what will come next.

@@

\subsubsection{Summary}

@@

>>>>

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See an example run-though of the videos here: }

%% PART UNFINISHED: link for example video

@@

\subsection{Results}
\label{sec:results1}

@@

\begin{table}[t]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\ 
    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\ 
    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\ 
    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\ 
    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\ 
    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\ 
    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\ 
    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\ 
%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
\label{tab:e1_look}
\end{table}

\begin{table}[h!]
\begin{center}
  \begin{tabular}{lcccl}
  \textbf{\textit{Children}} &&&& \\
    \hline
                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
    \hline
    (Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
    Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
    LgCond$=$\textit{non-English} &          -3.29347 &   0.96045 & -3.429 & 0.000606 *** \\
    Type$=$\textit{non-Question}          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
    Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  **\\
    Age*LgCond$=$\textit{non-English} &       0.52065 &   0.21190 &  2.457 & 0.014008 ** \\  
    Age*TypeS$=$\textit{non-Question}          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
    LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
    Type$=$\textit{non-Question} &&&& \\
    Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
    Type$=$\textit{non-Question} &&&& \\
    \hline
  &&&& \\
  \textbf{\textit{Adults}} &&&& \\
    \hline
                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
    \hline
    (Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
    LgCond$=$\textit{non-English}    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
    Type$=$\textit{non-Question}             &        -4.4953 &    1.3139 & -3.421 & 0.000623 *** \\
    Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
    LgCond$=$\textit{non-English}* &           3.2972 &    1.6101 &  2.048 & 0.040581 * \\ 
    Type$=$\textit{non-Question} &&&& \\
    LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
    Duration &&&& \\
    Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
    Duration &&&& \\
    LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
    Type$=$\textit{non-Question}* &&&& \\
    Duration &&&& \\
    \hline
  \end{tabular}
\end{center}
  \caption{Model output for children and adults' anticipatory gaze switches.}
\label{tab:E1-models}
\end{table}

@@

Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05) and a language condition-by-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05). There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{t}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{t}=-1.27, \textit{p}$=$.2, respectively).

@@

\subsubsection{Random baseline comparison}
\label{sec:randbaseline1}

We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows. This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; the analysis window is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), but this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}). Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each randomly permuted version of the data.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$300 msec around the inter-turn gap.} 
\label{fig:shuffling}
\end{figure}

This procedure effectively de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movements. The resulting anticipatory gazes from the randomly permuted analysis windows represent an average anticipatory gaze rate over all possible starting points: a random baseline. By running the real and randomly permuted data sets through identical statistical models, we can also estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.

We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data. For those interested, plots of the beta and standard error distributions are available in the Appendix.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test. For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p} $=$ .001). This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.

\begin{figure}[h!]
  \centering
  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
  \hfill
  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates. Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
\end{figure}

% PART UNFINISHED (potentially)
% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}). Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}). The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
\end{center}
\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data. Vertical bars represent the standard error.} 
\label{fig:E1-randvsreal}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
\end{center}
\caption{Pairwise comparisons of the language condition effect across ages for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).} 
\label{fig:E1-lgageinteraction}
\end{figure}

\paragraph{Developmental effects}
The model of the children's data revealed a significant interaction of age and language condition (tab:E1-models) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}). To further explore this effect, we compared the average effect of language condition across all ages: we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data. For each random permutation, we then made pairwise comparisons of the average difference scores across ages. Figure \ref{fig:E1-lgageinteraction} plots the real-data difference scores against the random-data difference score distribution for each pairwise age comparison, showing that 3- and 4-year olds were affected equally by language condition, but that 5-year-olds affected less than both 3- and 4-year-olds (with 99.52\% and 99.96\% of difference scores greater than the randomly permuted data, respectively, i.e., differences of \textit{p} $<$ .01 and \textit{p} $<$ .001).

\subsection{Discussion}
\label{sec:discussion1}

Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions (Table \ref{tab:e1_look} ; Figure \ref{fig:E1-randvsreal}). Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E1-ChiTs}).

Language condition (English vs. non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}). First, children made more anticipatory switches overall in English videos, compared to non-English videos. This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}. Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch when they can expect an immediate response from the addressee (i.e., an answer). In itself, this explanation also suggests that children are picking up on lexical cues to questionhood in anticipating upcoming answers (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).

Children's anticipatory gaze switches were also affected by their age, but only in the non-English videos: 3- and 4-year-olds made many more anticipatory switches when watching videos in English compared to non-English, but this effect of language condition had attenuated significantly by age 5 (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}; Figure \ref{fig:E1-lgageinteraction}). This interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues. Prior work on children's turn-structure anticipation proposed that children's turn-end predictions center on lexicosyntactic structure (and not, e.g., prosody) as they get older. The current results suggest more flexibility in children's predictions; when they do not have access to linguistic structure they are more likely to rely on alternative cues as they get older.

Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}). This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.

Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}). Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected. Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally significant compared to the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}). Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}). This interaction suggests that adults were less likely overall to make anticipatory switches at non-questions (as is evident for adults and children in Figure \ref{fig:E1-randvsreal}), and so did not benefit from extra time to do so compared to long gaps for questions.

\subsubsection{Summary}
\label{sec:summary1}

Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns that will have immediate responses from the addressee. Children's anticipatory gaze rates for questions and non-questions in English was stable across ages and comparable to adult behavior (Figure \ref{fig:E1-randvsreal}), suggesting that they can identify questions in native stimuli with adult-like competence by age three. Although participants' ability to recognize questions was facilitated by lexical access (i.e., English vs. non-English), the prosody in the non-English videos was non-native, and so the experimental design can not conclusively show which linguistic cues children relied on in the English videos to identify question turns.

Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants, especially adults, in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time. In the case of adults there was no overall effect of language condition, only an interaction between language condition and transition type: adults spontaneously anticipated upcoming speakers whether they had lexical access or not, but when given lexical access, focused their anticipations on turns with immediate responses (questions).

Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions. We had hypothesized that a projected response might affect participants' predictions but it in fact seems to be driving many of the effects in our data. Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}). Prior work using online, metalinguistic tasks has shows that participants can use linguistic cues to accurately predict upcoming turn ends. The current results suggest that, in their spontaneous predictions, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses. Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction. This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.

Overall, children and adults behaved relatively similarly and our language manipulation (English vs. non-English) was too coarse to comment on when children begin to use different types of native linguistic cues (e.g., prosody vs. lexicosyntax); we would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues. To see the emergence of anticipatory gaze switching we would also need to include younger children since participants already reliably made anticipatory gaze switches at age three. In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.

@@

@@

% PART UNFINISHED: APPENDIX
